Natural sources of uncertainty and their implications

### 2017-04-27 book_chapter/paper.tex
first lines for reference games

>>>>

Linguistic interaction, as captured by a sim-max game, has a rather narrow scope. There is a fixed prior and fixed utility function and a fixed similarity measure between states. If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t$. Here is a natural example of  a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue $t$; the closer Bubu's color choice matches Kiki's idea, the better. We could say that Kiki wants to describe the color for Bubu to pick out. It's a natural thing to want to do with language. But a lot of communication with vague and contiguously variable properties is rather different. In the general spirit of Wittgenstein's language games, we should look at some more cases of language use side by side. Let us therefore consider a continuum of cases of referential communication.

A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object. One crucial difference to descriptive communication and sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.

Interestingly, referential communication can take place under a plethora of different epistemic conditions. A lot of modeling work considers \emph{reference games without displacement} where sender and receiver communicate about a context, i.e. a set of objects, which is immediately accessible in their direct perceptual environment \parencite{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.\footnote{add more references later!} Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context. This, of course, need not be so. \textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}. Kiki asks Bubu to fetch the red book from the library. No other book than the one Kiki needs will serve. Yet when Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki. Yet other situation are imaginable. In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects. This is the set-up for the classic director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
